Bioconductor workflow for single-cell RNA sequencing: Normalization, dimensionality reduction, clustering, and lineage inference

### 2017-07-05 workflow/workflow.Rmd
clean up a bit

- solve loading packages issues (DR), DONE!
- versions of packages we want to use in develop branches and change functions in text accordingly (EP and DR), [DR: scone and zinbwave now OK!]
- update clustering section with new clusterExperiment functions (EP),
- add details in section preprocessing (DR), DONE,
- address Sandrine's comments in sections slingshot and de (KS),
- try running workflow with more genes (FP),
- plot gene variance density (FP),
- clean up singlecellworkflow package (FP),
- update NAMESPACE and DESCRIPTION (FP),
- install and run package using cache instead of saved files for zinbwave and clusterExperiment and making sure we can build the vignette using R 3.4 and bioc-devel 3.6 (all?),
- make singlecellworkflow package public (FP).]

@@

Note that in order to successfully run the workflow, we need the devel versions
of the Bioconductor packages `scone (>=1.1.2)`, `zinbwave (>=0.99.5)`, and `clusterExperiment (>=1.3.2)` [DR: Elizabeth, please check!]. We recommend running Bioconductor 3.6 (currently the devel version; see https://www.bioconductor.org/developers/how-to/useDevel/).

@@

```{r zinb}
print(system.time(se <- zinbwave(core, K = 40, X = "~ Batch",  
                                 residuals = TRUE,
                                 normalizedValues = TRUE)))
```

@@

[DR: Fanny, did you figure out why we get infinite values? If not, we need to at least explain why we are doing what we do below.]
[FP: I created a pull request on github. Let me know what you think.]

```{r norm}
norm <- assays(se)$normalizedValues
if (sum(is.infinite(norm))>0){
  maxNorm = max(norm[!is.infinite(norm)])
  assays(se)$normalizedValues[is.infinite(norm)] <- maxNorm
  norm <- assays(se)$normalizedValues
}
norm[1:3,1:3]
```

@@

```{r barplotOurs, fig.cap="Barplot of number of cells per cluster, for our workflow's clusterint."}
plotBarplot(ceObj) 
```

[FP: Should we add here the short paragraph Elizabeth was talking about to compare our found clusters versus the original published clusters?]

```{r addPublishedClusters, fig.cap="Barplot of number of cells per cluster, for original published clustering."}
ceObj<-addClusters(ceObj, colData(ceObj)$publishedClusters, 
                   clusterLabel = "publishedClusters")
plotBarplot(ceObj,whichClusters=c("combineMany","publishedClusters")) 
```

EP: the label `clusterLabels` for the original labels might do better to be changed to `publishedClusters` so that it is more informative. Also, we don't do a direct comparison of the `publishedClusters`
 to the found clusters -- is that relevant? If so, it is probably better to add the published clusters via `addClusters` so it is easier to treat them like clusters. This is perhaps useful for a workflow since this is a common task that `clusterExperiment` makes easy. Here is code that does this:

[FP: after the modifications, do we still need function addCluters? If we want to illustrate that clusterExperiment package has a convenient function to add clusters to a clusterExperiment object, we can keep it and say that here it is not necessary.]

@@

[EP: If we do the above code, then the `plotHeatmap` command below should be modified to have the `publishedClusters` under the `whichClusters` argument. ]

[FP: Elizabeth, I changed the `clusterLabels` to `publishedClusters`, more informative indeed. Maybe, we could add the plotClusters to compare the published and our own clusters and the plotBarplot to illustrate our previous call to function table()? I modified the code accordingly. If you don't like it, feel free to change, it is just a proposition.]

>>>>

- make sure versions package ok and pushed to bioc (DR, EP, FP),
- update clustering section with new clusterExperiment functions (EP),
- address Sandrine's comments in sections slingshot and de (KS),
- try running workflow with more genes (DR),
- clean up singlecellworkflow package (FP),
- update NAMESPACE and DESCRIPTION (FP),
- install and run package using cache instead of saved files for zinbwave and clusterExperiment and making sure we can build the vignette using R 3.4 and bioc-devel 3.6 (all?),
- make singlecellworkflow package public (FP).]

@@

Note that in order to successfully run the workflow, we need the devel versions
of the Bioconductor packages `scone (>=1.1.2)`, `zinbwave (>=0.99.6)`, and `clusterExperiment (>=1.3.2)` [DR: Elizabeth, please check!]. We recommend running Bioconductor 3.6 (currently the devel version; see https://www.bioconductor.org/developers/how-to/useDevel/).

@@

```{r zinb}
print(system.time(se <- zinbwave(core, K = 50, X = "~ Batch",  
                                 residuals = TRUE,
                                 normalizedValues = TRUE)))
```

@@

```{r norm}
norm <- assays(se)$normalizedValues
stopifnot(sum(is.infinite(norm))==0)
norm[1:3,1:3]
```

@@

```{r barplotOurs, fig.cap="Barplot of number of cells per cluster, for our workflow's clusterint."}
plotBarplot(ceObj, legend = F) 
```

[FP: Should we add here the short paragraph Elizabeth was talking about to compare our found clusters versus the original published clusters?]
[FP: we should add a comment on cluster 7.]

```{r addPublishedClusters, fig.cap="Barplot of number of cells per cluster, for original published clustering."}
ceObj<-addClusters(ceObj, colData(ceObj)$publishedClusters, 
                   clusterLabel = "publishedClusters")
plotBarplot(ceObj,whichClusters=c("combineMany","publishedClusters"), legend = F) 
```

@@
